Think deep for reef resilience

### 2025-08-13 paper/paper_draft.qmd
Reorder methods sections
Add initial bayesian regression methods paragraph and update log info

@@

Reefs with greater median depth levels maintain coral cover above indicative carbonate
budget levels for a longer time frame compared to reefs with shallower median depths
[@fig-access-esm1-5-depth-carbonate]. Reef median depth has a moderate positive correlation
with the number of years a reef maintains coral cover above the required threshold
[@fig-access-esm1-5-depth-carbonate]. Depth - carbonate budget maintainence correlation
decreases as $\beta_{0}$ increases. This is likely due to an increasing number of reefs
being unable to maintain coral cover above threshold levels as the threshold increases, with
values becoming increasingly right-skewed. The natural log of total connectivity strength
was found to have no correlation with the number of years a reef maintains cover above the
required thresholds [@sfig-access-esm1-5-conn-carbonate]. Log connectivity strength -
carbonate budget maintainence correlation slightly decreases as the required carbonate
budget threshold increases.

@@

```{julia}
#| echo: false
#| warning: false
#| message: false
#| results: asis
#| label: tbl-access-esm1-5-bioregion-regression
#| tbl-cap: Multinomial logistic regression coefficients for the statistical model containing median reef depth, natural-log total connectivity strength, natural-log source-to-sink ratio and ACCESS-ESM1-5 mean timeseries DHW value as predictor variables. ACCESS-ESM1-5 bioregion scale reef cluster assignment is the response variable with three levels, 'low' (regression reference level), 'medium' and 'high'.

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef natural log of
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Colour of each point represents the natural log of scaled reef total connectivity
strength. Displays results from median timeseries using
ACCESS-ESM1-5.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#sfig-access-esm1-5-conn-carbonate
fig-align="center" width="14.82cm"}

@@

![Source-to-sink connectivity ratios (in natural log scale) for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median source-to-sink ratio for each cluster and grey
points represent the ratio value for each reef. Vertical grey lines represent a
source-to-sink ratio of 0 (in natural log scale) indicating equal outgoing and incoming
connectivity levels. Displays results for ACCESS-ESM1-5
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi
fig-align="center" width="14.82cm"}

![Total connectivity strength (in natural log scale) for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median natural log of total connectivity strength for
each cluster. Displays results for ACCESS-ESM1-5
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-totalconnectivity
fig-align="center" width="14.82cm"}

>>>>

To quantify the influence of reef characteristics on cluster assignment we used a Bayesian
ordinal mixed regression approach. Using Turing.jl [@fjelde2025] we model the influence of
five predictors, median depth, log total connectivity strength, log source to sink ratio,
mean DHW, log reef area and mean DHW tolerance level on cluster assignment, including random
intercepts for levels of **bioregion** and GCM identity.

@@

Reefs with greater median depth levels maintain coral cover above indicative carbonate
budget levels for a longer time frame compared to reefs with shallower median depths
[@fig-access-esm1-5-depth-carbonate]. Reef median depth has a moderate positive correlation
with the number of years a reef maintains coral cover above the required threshold
[@fig-access-esm1-5-depth-carbonate]. Depth - carbonate budget maintainence correlation
decreases as $\beta_{0}$ increases. This is likely due to an increasing number of reefs
being unable to maintain coral cover above threshold levels as the threshold increases, with
values becoming increasingly right-skewed. The base-10 log of total connectivity strength
was found to have no correlation with the number of years a reef maintains cover above the
required thresholds [@sfig-access-esm1-5-conn-carbonate]. Log connectivity strength -
carbonate budget maintainence correlation slightly decreases as the required carbonate
budget threshold increases.

@@

```{julia}
#| echo: false
#| warning: false
#| message: false
#| results: asis
#| label: tbl-access-esm1-5-bioregion-regression
#| tbl-cap: Multinomial logistic regression coefficients for the statistical model containing median reef depth, base-10-log total connectivity strength, base-10-log source-to-sink ratio and ACCESS-ESM1-5 mean timeseries DHW value as predictor variables. ACCESS-ESM1-5 bioregion scale reef cluster assignment is the response variable with three levels, 'low' (regression reference level), 'medium' and 'high'.

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef base-10 log of
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Colour of each point represents the base-10 log of scaled reef total connectivity
strength. Displays results from median timeseries using
ACCESS-ESM1-5.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#sfig-access-esm1-5-conn-carbonate
fig-align="center" width="14.82cm"}

@@

![Source-to-sink connectivity ratios (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median source-to-sink ratio for each cluster and grey
points represent the ratio value for each reef. Vertical grey lines represent a
source-to-sink ratio of 0 (in base-10 log scale) indicating equal outgoing and incoming
connectivity levels. Displays results for ACCESS-ESM1-5
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi
fig-align="center" width="14.82cm"}

![Total connectivity strength (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median base-10 log of total connectivity strength for
each cluster. Displays results for ACCESS-ESM1-5
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-totalconnectivity
fig-align="center" width="14.82cm"}
